Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex

\title{{\normalfont Response to reviewers of the manuscript \texttt{2016GL070261} }
       ``Seasonality of submesoscale
       dynamics \\in the Kuroshio Extension''}
\author{Cesar B. Rocha,
        Sarah T. Gille, Teresa K. Chereskin, \& Dimitris Menemenlis}
\date{}
\begin{document}

%\include{symbols}

\maketitle

\section*{Summary}

We thank the reviewers for the speedy reviewer, constructive criticism, and
great suggestions.

@@

          \bdp
              We agree that the planning of SWOT is perhaps the most immediate
              application of our results. We have added a paragraph dedicated to SWOT (new
              paragraph 2 of section 1). We have also added a discussion about the
              implication for SWOT in section 6. Thanks for making us provide a better
              context for our study.
          \edp

  \item {\it There needs to be a discussion about the effective resolution of your model. The fact that
        the model has a resolution of dx = 2.3 km and 4.6 km does not mean it can resolve oceanic
        phenomena at this scale. Its effective horizontal resolution is closer to 8 times the grid
        resolution (Soufflet et al. 2013). This corresponds to 18 km and 35 km for LLC4320 and
        LLC2160 output, respectively. [A more exact value could be obtained by estimating zonal
        and meridional spectra (of u, v or SSH) and identifying the high wavenumbers / low
        wavelengths at which the spectral energy falls precipitously.] This means the model is not
        resolving a number of submesoscale coherent vortices (SCVs) that may occur as a result of
        ML baroclinic instability; these have diameters close to 5 km. So the follow-up questions
        from a dynamical point-of-view become (1) where does the unresolved energy come from or
        go to? and (2) how does this affect your results? These questions need to be addressed or at
        least discussed if the main topic of your paper is submesoscale dynamics.}\\

        \bdp
            We are aware of the this issue --- the effective resolution based on the KE spectrum,
            was already mentioned
            in the supporting information to the first draft.  In
            fact, the smallest scale discussed in our paper (10 km) was set by the effective
            resolution of highest-resolution simulation. To be more explicit, we have now included
            the effective resolution together with the nominal resolutions in the main text (Section 2).\\

            The ``unresolved energy'' does not come from anywhere; it is unresolved.
            To the extend that 1-km-scale are associated with a forward
            energy transfer, then their effects on the scales concerning our study
            (10-100 km) are likely energetically insignificant.
            Also important to this discussion is that higher resolution simulations will resolve
            1-km-scale eddies and high-wavenumber IGWs that are generated through wave-wave
            interactions and wave-vorticity interactions.\\

            Buckingham et al. (2016) have a very good point that models and observations
            to date may not resolve the mixed-layer deformation scale in summer; this is no
            different for our simulations. But summer is also the season when lateral buoyancy
            gradients (and thus available potential energy) are minimum within the mixed layer,
            so the reviewer's conjecture that this is the main source of seasonally varying eddy
            kinetic energy is not obvious.\\

            For geostrophically balanced mixed-layer instabilities (which we call
            shallow baroclinic instabilities), linear theory
            predict maximum growth rates at about $2.6\times R_{ml}$, where
            $R_{ml}$ is the mixed-layer deformation scale. But in equilibrated
            QG turbulence driven by ML instabilities (Callies et al. 2016) the
            deformation scale has only a catalytic role, where eddy baroclinic
            energy is converted into (mixed-layer) eddy barotropic energy; as
            in interior baroclinic geostrophic turbulence, most of the baroclinic
            conversion occurs at larger scales (Callies et al. 2016).

            Still, we recognize that this issue may be further investigated and
            we included a paragraph to the discussion.

        \edp

@@

        One way to define mesoscale and submesoscale is to define what scales of motion a
        particular sensor will resolve. While this would not be optimal, it is objective. This is
        probably the motivation for the definition used by Callies and Ferrari (2013), since the
        swath altimeter (as initially proposed) would resolve scales greater than 1 km but
        smaller than the present-day, profiling satellite altimeters – i.e., eddy-like features
        with diameters greater than 80-100 km (see Chelton et al. 2011, Progress).
        Unfortunately, this definition has the adverse effect of changing for each sensor.

        A better definition of the submesoscale might be founded on dynamical arguments.
        Wunsch and Stammer (1998) introduce the geostrophic equations and note that the
        lower limit for which this approximation is strictly valid is ~30 km. This conveniently
        defines the lower limit of the mesoscale regime. I would therefore define the
        submesoscale regime as anything smaller than this but one for which the hydrostatic
        approximation remains valid (e.g., 1 km).

      The motivation for wanting to use the term ``submesoscale'' in the title is clear –
      submesoscale motions are a hot topic within the community and it catches individuals’
      eyes. This is fine. But what must be made clear within the manuscript is what scales of
      motion and what force balances (i.e., dynamical regimes) are being described.}\\

      \bdp
          We respect the reviewer's opinion about the definition of submesoscales.
          But there seems to be no clear definition in the literature --- the
          Stammer $\&$ Wunsch paper states that their
          definition of mesoscales is ``very rough'' and ``there is no generally
          agreed upon definition''. There is also irritation that oceanographers
          decided to term mesoscales the flows with dynamics analogous to
          the meteorological synoptic scale motions.\\

@@

          Some investigators define mesoscales as flows with scales near the
          1st baroclinic deformation radius. But there is significant confusion
          owing to factors of  $2 \pi$ --- e.g., some investigators would use $R_1 = 30$ km,
          others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
          (Wunsch \& Ferrari 2004; Ferrari \& Wunsch 2009).
          To add yet another controversy, Larichev \& Held 1995 argue
          that the deformation length is not the panacea as linear theorists
          suggest. In baroclinic geostrophic turbulence, the bulk of the
          energy production occurs on scales much larger than the deformation
          radius; the energy-containing scales of equilibrated baroclinic geostrophic
          turbulence are also larger than the deformation radius (e.g., Larichev \& Held 1995).\\

          According
          to our definition the mesoscales are the energy-containing scales of the
          flow --- the eddies that contain most of the eddy kinetic energy in the ocean
          (Ferrari \& Wunsch 2009). On observational
          grounds, this scale varies roughly from 100-300 km (from the centroid of KE wavenumber
          spectrum from current altimeters for example). We then \textit{define}
          the energetically subdominant scales, the scales smaller than about 100 km.
          Our lower wavelength definition is limited by the effective resolution
          of our higher resolution simulation ($\sim$10 km).
          This definition is consistent with Callies \& Ferrari 2013 (they did not
          define mesoscales/submesoscales based on scales that a sensor resolves
          as the reviewer suggests).\\

          We understand that our definition is imperfect --- we may amuse some readers but
          irritate others. But
          besides the unsettled definitions in our community and consistency with our recent
          work (Rocha et al. 2016), there are at least two reasons for sticking to
          our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
              2008); incidentally current altimeters do not resolve
              scales smaller then about 100 km; and (ii) our goal is to determine the flows
              (geostrophically balanced turbulence, inertia-gravity waves, etc) that
              dominate the 10-100 km horizontal
              scales and their seasonality. In any event,  readers that make to the first sentence of our
              abstract will find our explicit definition of submesoscales.
      \edp

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex

        \bdp
          We have added both nominal and effective resolutions.
        \edp

  \item {\it (Section 1, Paragraph 3) What types of horizontal and temporal scales for IGWs are
        we talking about here? How would this fall onto the Garrett-Munk (GM) internal
        wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
        flows, not just internal waves, despite that the name suggests this.)}\\

        \bdp
          We explicit state that those are submesoscale IGWs. Using
          our definition of submesoscales, these are IGWs
          with horizontal scales about 10-100 km\footnote{Bill Young suggested the
          term submesoscale IGWs.}; IGWs have periods  The GM spectrum is not accurate
          close to the surface (most of our results concern surface currents and SSH,
          and it does not account for tidal and inertial peaks. We believe that the
          new plot of the wavenumber-frequency spectrum of the SSH
          variance (new Figure 4) --- following the reviewer's suggestion ---
          will more clearly convey the idea that there is no horizontal scale
          separation between submesoscale IGWs and submesoscale turbulence.\\

          If the reviewer is using ``balanced'' as a short to ``geostrophically
          and hydrostatically 
          balanced'', then the comment about the GM spectrum containing both balanced
          and unbalanced flows is incorrect.  GM used linear theory
          of IGW to synthesize observations, mostly mooring data --- they
          fitted an analytic model to (super-inertial) frequency spectra and
          different vertical modes (see Walter Munk's chapter
          in EPO 1981 and references therein).
          One could argue
          that other flows project onto similar temporal scales, but those motions
          are super-inertial (e.g., stratified turbulence), not (sub-inertial)
          balanced flows.
        \edp

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex

%\include{symbols}

\maketitle

\section*{Summary}

In this revised version, we have addressed the reviewers
comments. The most notable change is the inclusion of a wavenumber-frequency
spectral analysis suggested by reviewer 2. This analysis helps us vindicate
our claim that inertia-gravity waves, particularly super-inertial waves,
account for most of the summertime SSH variance and surface KE at scales
between 10 and 100 km. We now also place our study in the context of SWOT and discuss
the implications of our findings for the estimation of geostrophic velocity from
high-resolution altimeters. We thank the reviewers  for the speedy review, constructive
criticism, and fine suggestions, which helped us improve our discussion and
provide a better context for our study.

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex

We changed jPDF with joint-PDF; it does look much better.

\item {\it (Section 4, last paragraph) ``...whereas divergence is moderately, negatively skewed as
        predicted by ...'' Specifically mention convergence/downwelling.}

\item {\it (Figure 1) ** Place panels (d) and (e) before panels (b) and (c) since this is the order
        in which you refer to them in the manuscript. ** Difficult to see contours in (d) and
        (e). ** Consider overlaying mixed layer depth on these transects since it is mentioned
        in the text.}

        \bdp
          We have changed the order of the panels and plotted the mixed layer depth
          on the density plots.
        \edp

### 2016-08-23 writeup/rocha_etal.tex
Adds new figure 1

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
as deep as 300 m (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
than 40 m (Figure \ref{fig1}c).
Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supporting information).

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.7\textwidth]{figs/fig1_2.pdf}
 \caption{\small (a) The study region with the subregion where the LLC outputs are
          analyzed. Colors represent the topography and white lines are contours of absolute
          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of
          transects
          of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e) . The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

>>>>

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
than 40 m (Figure \ref{fig1}c).
Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supporting information).

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.7\textwidth]{figs/fig1_2.pdf}
 \caption{\small (a) The study region with the subregion where the LLC outputs are
          analyzed. Colors represent the topography and white lines are contours of absolute
          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of
          transects
          of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e).
          In (b) and (c) the white line represents the mixed layer depth. The snapshots were
          taken at 12:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}
